Cross-sectoral conservation of the Indian Ocean

### 2022-03-02 Msc_Thesis_Lea_v2.Rmd
R file because bugs in .rmd file
All code from .rmd + area calculation chunk for PMN and CFC in Indian Ocean

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures? Or is it enough to have columns next to PUs? RESOLVED: yes, stacking = creating columns next to PUs.
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
+ shipping raster : cells have certain values (# of ships recorded?? I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not? RESOLVED: check 2008 metada, run prio problem with only shipping
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see). RESOLVED: yes, use fx from Jase & Tin. Change scale to remove outliers/crop to planning region.
+ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html. RESOLVED: cut up chuncks of code with separate ones for plots
+ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. RESOLVED: use st_contains after crs = 4326 & st_transform(cCRS)
+ add list of species as cons_feats: EN, VU etc (Ant's email to mining colleague)? RESOLVED: fine-tune later
+ shipping raster (Halpern et al., 2013) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution <1°. Resolution = 934m*934m (! 1° at equator = 111km = 60 nm; [convert](https://www.johndcook.com/how_big_is_a_degree.html); resolution too fine so need resampling). RESOLVED: no resampling, overlay raster with PUs (exactextractr) to attribute mean raster cells value to PUs (no weighted mean because raster cells are all equal-area because Mollweide proj)
+ how to remove EEZs from PUs? RESOLVED: find EEZ shp and spatial predicates, inverse overlay with Bndry, recreate PUs
+ setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)

@@

+ ongoing - pricing deep-sea mineral resources (I am looking for more papers especially regarding CFC)
+ check research questions
+ overlay shipping & PUs + plot with log10?
+ cross inactive vents & PUs - done
+ how to assign price to mining shps
+ assigning targets - write chunk

@@

@@

PU_size <- 1000 # in km2, but check resolution for this project?
Shape <- "Hexagon" # shape of PUs
MinDepth <- 0
MaxDepth <- 200 # but check for deep-sea?
PUs <- fSpatPlan_Get_PlanningUnits(Bndry, world, PU_size, Shape)

@@

```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)

@@

@@

```{r Shipping plot, echo = FALSE}
Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
plot(Shipping, main = 'Shipping routes')

#ShippingRobin <- terra::
#exactextractr::exact_extract(Shipping, PUs, mean)
```

@@

@@

Could set cut-off for cost outliers at 15,000 USD or 40,000 USD or no cut off - depending on whether we want to favour variation or follow standard procedures for outliers (see histograms and maps).

@@

  FishingCost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
    st_as_sf() %>% # Convert to sf
    st_transform(cCRS) %>% # Transform to robinson
    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
    rename(FishingCost = layer)
}

FishingCost <- fSpatPlan_Get_Cost(PUs, cCRS)

hist(FishingCost$FishingCost) #max value >200,000
FishingCostPlusOne <- FishingCost$FishingCost + 1
hist(log10(FishingCostPlusOne)) #drop after log10(x)=4, x=10^4=10,000?

@@

```
FishingCostOutliers <- filter(FishingCost, Cost >= 22000) # check outliers

@@

```
p_ConsFeats <- problem(PUs, Cons_feats$Features, cost_column = "NoCost") %>%
  add_min_set_objective() %>%
  add_relative_targets(Cons_Feats$Targets) %>% #
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%

````

@@

```
p_ConsFeat_Fishing <- problem(PUs, features = ConsFeatures, cost_column = "FishingCost") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
```

@@

@@

```
# create problem with added locked out constraints using spatial polygon data
locked_out <- sim_pu_polygons[sim_pu_polygons$locked_out == 1, ]
p5 <- p1 %>% add_locked_out_constraints(locked_out)
```

@@

p_ConsFeat_Mining <- problem(PUs, features = ConsFeatures, cost_column = "PMNCost") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

@@

p_ConsFeat_MiningFishing <- problem(PUs, features = ConsFeatures, cost_column = "MiningFishing") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

@@

*Same as p_ConsFeat_MiningFishing but with shipping routes locked out as well*

@@

p_ConsFeat_Shipping <- problem(PUs, features = ConsFeatures, cost_column = "MiningFishing") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(ConsFeatures$amount) %>% # representation targets (Area)
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  add_locked_out_constraints("Shipping","Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

>>>>

+ how to group all files (ex: all EBSAs + IBAs + IMMAs + vents + seamounts...) into one, to create ConsFeatures? Or is it enough to have columns next to PUs? RESOLVED: yes, stacking = creating columns next to PUs.
+ should we use Sharma (2011) + CRU Consulting (2019) to assign USD values to PMN areas? see PMN subsection. RESOLVED, yes, can use for this project (need to cross-check if want to publish)
+ shipping raster : cells have certain values (# of ships recorded?? I can't figure from the attached files, metadata is confusing), could allow to infer USD value of route (# of ships * approx cost of fuel for average cargo to go through 1 PU, which is 1000sqkm)? So we wouldn't try to compute the cost of deviating from any given route, but rather compute the price for each ship to go through 1 PU * the number of ships going through that PU. It sounds fairly logical to me, but maybe it's not? RESOLVED: check 2008 metada, run prio problem with only shipping
+ fishing cost layer : there is a ready-made fx in SpatPlan repo (thanks Jase & Tin), should we use it for this project? see fishing subsection (but issues with plotting  as you can see). RESOLVED: yes, use fx from Jase & Tin. Change scale to remove outliers/crop to planning region.
+ I'm using echo = FALSE, message = FALSE, warning = FALSE, results = "hide" and still can't get rid of some coding output in the .html. RESOLVED: cut up chuncks of code with separate ones for plots
+ I st_to_sf the vents from .csv with PUs' crs but have some issues intersecting the points with polygs, will look more into it. RESOLVED: use st_contains after crs = 4326 & st_transform(cCRS)
+ add list of species as cons_feats: EN, VU etc (Ant's email to mining colleague)? RESOLVED: fine-tune later
+ shipping raster (Halpern et al., 2015) is too heavy to be converted to polygon, what could be other methods (not using as.polygon)? RESAMPLE if resolution <1°. Resolution = 934m*934m (! 1° at equator = 111km = 60 nm; [convert](https://www.johndcook.com/how_big_is_a_degree.html); resolution too fine so need resampling). RESOLVED: no resampling, overlay raster with PUs (exactextractr) to attribute mean raster cells value to PUs (no weighted mean because raster cells are all equal-area because Mollweide proj)
+ how to remove EEZs from PUs? RESOLVED: find EEZ shp and spatial predicates, inverse overlay with Bndry, recreate PUs
+ setting  targets. RESOLVED: 30% for rpz area, 70% for important/rare feature; create dataframe with 1 column: features, 1 column: targets (match order)
+ cross inactive vents & PUs - done

@@

+ ongoing - pricing deep-sea mineral resources (I am looking for more papers especially regarding CFC)
+ check research questions
+ overlay shipping & PUs + plot with log10?
+ how to assign price to mining shps / create cost layer?
+ assigning targets - Cons_feats tibble created

@@

# define region and boundary

@@

# intersect bndry and high seas

ABNJ <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/World_High_Seas_v1/High_Seas_v1.shp") 
ABNJRobin <- st_transform(ABNJ, cCRS) # high seas in robin proj, does not work when using pipe
ABNJRobBuff  <- st_buffer(ABNJRobin, 0) # need to create buffer of 0 to avoid Error in CPL_geos_op2(op, x, y) : Evaluation error: TopologyException: Input geom 0 is invalid: Self-intersection at or near point -> overwrite offending geometries
BndryABNJ <- st_intersection(ABNJRobBuff, Bndry)

#create PUs

PU_size <- 1000 # in km2, but check resolution for this project?
Shape <- "Hexagon" # shape of PUs
MinDepth <- 0
MaxDepth <- 200 # but check for deep-sea?
PUs <- fSpatPlan_Get_PlanningUnits(BndryABNJ, world, PU_size, Shape)

### method with EEZ took forever and was not created the PUs I wanted, so I chose to keep high seas shp instead of using EEZ shp
#EEZ <- st_read ("~/Documents/MscThesis/GitHub/SpatialPlanning/Data/World_EEZ_v11/eez_v11.shp") 
#EEZRobin <-  st_transform(EEZ, cCRS)
#BndryABNJ <- st_difference(Bndry, EEZRobin)
#PUs <- fSpatPlan_Get_PlanningUnits(Bndry, world, PU_size, Shape) # error : st_crs(x) == st_crs(y) is not TRUE ?
```

@@

```{r vents plot, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}
world <- ne_countries(scale = "medium", returnclass = "sf")
IndianO <- c(xmin = 17.47941, ymin = -38.7402, xmax = 129.552, ymax = 33.82872)

@@

```{r Conservation features and targets, echo = FALSE, message = FALSE, warning = FALSE}

@@

```{r Shipping plot, echo = TRUE}

Shipping <- terra::rast("~/Documents/MscThesis/GitHub/SpatialPlanning/Shp/Costs/Shipping/shipping.tif")
#View(Shipping)
#plot(Shipping, main = 'Shipping routes')
ShippingRobin <- terra::project(Shipping, "ESRI:54030") #transform to robin proj
#plot(ShippingRobin, main = 'Shipping routes')
PUxShipping <- exactextractr::exact_extract(ShippingRobin, PUs, "mean") # attribute mean value of pixels in each PU to PU
PUs <- PUs %>% 
  mutate(ShippingIntensity = PUxShipping) # add column to PU data frame

#ShippingPlusOne <- Shipping$shipping +1
#ShippingLog <- log10(ShippingPlusOne)
#hist(log10(ShippingPlusOne))  #Warning: [hist] a sample of0% of the cells was used (of which 21% was NA); meaning +1 didn't work?

#ShippingRobin <- terra::project(Shipping, "cCRS") #Error: [project] cannot get output boundaries

#ShippingRobin <- project(Shipping, "+proj=robin +lon_0=0 +x_0=0 +y_0=0 #+datum=WGS84 +units=m +no_defs") # no error but does not assign projection

#ShippingRobin <- project(Shipping, "ESRI:54030", method = "near") #transform to robin proj # error about method "near"

@@

```
# not using this method because file to heavy to use as.polygons

@@

@@

  FishingCost <- terra::rast("Data/Cost/Cost_Raster_Sum.grd") %>% 
    terra::as.polygons(trunc = FALSE, dissolve = FALSE, na.rm=FALSE) %>% # Convert to polygon data
    st_as_sf() %>% # Convert to sf
    st_transform(cCRS) %>% # transform to robinson
    st_interpolate_aw(PUs, extensive = FALSE) %>% ## intersect with PUs
    rename(FishingCost = layer)
}

FishingCost <- fSpatPlan_Get_Cost(PUs, cCRS)
PUs <- PUs %>% 
  mutate(Fishing = FishingCost) #issues: no error but cannot check PU dataframe after this: Error in `[.data.frame`(x, i) : undefined columns selected; probably need to restart session
hist(FishingCost$FishingCost) 
#FishingCostPlusOne <- FishingCost$FishingCost + 1 #not needed, min >0 for PU=ABNJ
hist(log10(FishingCost$FishingCost)) #drop after log10(x)=4, x=10^4=10,000? #no longer needed since in ABNJ max val = ##USD 5000, and most values are 10^1 - 10^2 USD

@@

#section not needed anymore since no outliers when PUs=ABNJ

FishingCostOutliers <- filter(FishingCost, Cost >= 22000) # check outliers

@@

p_ConsFeats <- problem(PUs, features = Cons_feats$Features) %>% #issues because no cost_column
  add_min_set_objective() %>%
  add_relative_targets(Cons_feats$Targets) %>% 
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE)

@@

*code based on Alvise's work, not as code here to avoid issues when knitting*

p_ConsFeat_Fishing <- problem(PUs, features = Cons_feats$Features, cost_column = PUs$FishingCost) %>% 
  add_min_set_objective() %>% 
  add_relative_targets(Cons_feats$Targets) %>% 
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE)

@@

*code based on Alvise's work, not as code here to avoid issues when knitting*

@@

# create problem with added locked out constraints using spatial polygon data

```
use:

locked_out <- sim_pu_polygons[sim_pu_polygons$locked_out == 1, ]
p5 <- p1 %>% add_locked_out_constraints(locked_out)
```
or try

@@

p_ConsFeat_Mining <- problem(PUs, features = Cons_feats$Features, cost_column = PUs$PMNCost) %>% 
  add_min_set_objective() %>% 
  add_relative_targets(Cons_feats$Targets) %>% 
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) %>%
  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

@@

p_ConsFeat_MiningFishing <- problem(PUs, features = Cons_feats, cost_column = PUs$MiningFishing) %>% 
  add_min_set_objective() %>% 
  add_relative_targets(Cons_features$Targets) %>%
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) %>%
  add_locked_out_constraints("Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```

@@

@@

*here, Shipping = mean values of pixels overlaid on PUs, so intensity of shippings; not USD*

```{r Shipping problem, echo = TRUE, warning = FALSE}

p_ConsFeat_Shipping <- problem(PUs, features = Cons_feats$Features, cost_column = "PUs$Shipping") %>% 
  add_min_set_objective() %>% 
  add_relative_targets(Cons_feats$Targets) %>% 
  add_binary_decisions() %>%
  add_lpsymphony_solver(verbose = FALSE) #%>%
  #add_locked_out_constraints("Shipping","Expl_CFC", "Expl_PMN", "Expl_PMS", "Res_PMN", "Res_CFC")
```
